Numerical investigation of viscous flow instabilities through heterogeneous porous media. Advantages and limitations of unstructured and adaptive meshes in 2D and 3D domains under different mobility scenarios.

### 2017-08-11 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
Updates on Section 4.5

The last of the simulations introduced is a $3$D domain that will showcase the idea of the preferential flow path and finger growth using different permeability regions, as these are decribed in figure \ref{fig:3DChannel_PermMesh}. In large scale geological formations preferential flow allows much faster contaminant transport and creates significant consequences for ground-water quality.

\medskip
As already mentioned in section \ref{section:ViscousInstabilities}, \textit{viscous fingering} is characterized by the formation of flow patterns, called viscous fingers. The geometrical evolution of these patterns provides good indications on capturing and tracking any instabilities in the flow front. Figure \ref{fig:3DChannel_PermMesh} describes the permeability pattern as this is implied in a $3D$ domain. The surface on the left hand side of the domain is the area where fluid $1$ is injected with a velovity $ u=2.5 cm/sec $. No-slip conditions are implied everywhere else (\ie insulation, no exchange of material or Neumann boundary condition for mass transfer) while at the right hand side of the domain there is the outlet. The control volume is constructed in such a way so as to be able to focus and capture the preferential path of flow as this can be seen from figures \ref{fig:3DChannel_planes}-\ref{fig:3DChannel_Isosurf}.

>>>>

In this section, 3D simulations are introduced to investigate preferential flow pathways and finger growth through semi-pervious and impervious layers, as decribed in Figure \ref{fig:3DChannel_PermMesh}. In large scale geological formations preferential flow may allow fast contaminant transport and creates significant consequences for ground-water quality, and low recovery rate in hydrocarbon production fields.

\medskip
As mentioned in Section \ref{section:ViscousInstabilities}, \textit{viscous fingering} is characterized by the formation of flow patterns, called viscous fingers. The geometrical evolution of these patterns provides good indications on capturing and tracking any instabilities in the flow front. Figure \ref{fig:3DChannel_PermMesh} shows the initial permeability distribution for the simulations described in this section. The surface on the left hand side of the domain is the area where fluid $1$ is injected with a uniform velovity of 2.5 cm.s$^{-1}$. No-slip conditions are implied everywhere else (\ie no flux across the boundaries) except at the right hand side of the domain where the fluid is allowed to leave the domain. The control volume is constructed in such a way so as to be able to focus and capture the preferential path of flow as this can be seen from Figures \ref{fig:3DChannel_planes}-\ref{fig:3DChannel_Isosurf}.
